Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-06-04 limits/ms.tex
Addressed major comment and partially the medium one.

In this paper, we use a new program -- \Glass\ (Gravitational Lensing AnalysiS
Software) -- to determine what combination of lensing, stellar mass, and
stellar kinematic constraints best constrain the projected mass profile and
shape of several gravitational lens produced from mock data.  The
compute-intensive core of \Glass\ is a parallel implementation of the
high-dimensional sampling algorithm from \cite{2012MNRAS.425.3077L}.  Around
that core, \Glass\ has extensible code (in software terminology, a framework)
in Python, encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory}, we review the key
elements of lensing theory, stellar population synthesis, and stellar dynamics
we will need. In \secref{sec:mockdata}, we describe our mock data. In
\secref{sec:results}, we present our results from applying \Glass\ to these
mock data. Finally, in \secref{sec:conclusions} we present our conclusions.

@@

Unfortunately, the lens equation and the arrival times alone are not enough to
form a closed volume in the solution space. We have not imposed any strong
constraints on the density even so much as to ensure that it is non-negative
everywhere. We must therefore provide additional linear constraints so that we
only sample physical models. The explicit implementation for these priors has
been explained in \cite{2008ApJ...679...17C}, but can be summarised as
%
\begin{enumerate}
\item The density must be non-negative everywhere.
\item The density profile must have a slope everywhere $\le 0$.
\item The local gradient everywhere must point within $45^{\circ}$ of the center.
\item Density variations must be smooth.
\item Image parity is enforced.
\item The density is inversion symmetric. (Optional)
\end{enumerate}

Prior (ii) is necessary to prevent rings from forming where the inner region is
less dense. This would be unusual for a galaxy or a cluster. Prior (vi)
enforces a radial symmetry that requires radially opposed pixels (i.e., on
opposite sides of the center) to be equal. This prior is particularly important
for modelling doubles which must be symmetric. If a lens is suspected to be
symmetric this can also significantly improve the modelling time for quads.

In the simplest form, a single model for a lens is a tuple $\M = (\vec\kappa,
\vec\beta, \gamma_1, \gamma_2)$. A single model represents a single point in
the solution space polytope. Using the MCMC sampling strategy described in
\cite{2012MNRAS.425.3077L} we uniformly sample this space. Collectively, the
sampled models are referred to as an ensemble $\E = \{\M_i\}$, where we usually
generate $|\E| \sim 1000$ models.
\hilight{Referee suggests adding some timing information here. How long does it
take to generate 1000 models?}

@@

\figref{main results} and \figref{main results pixel-wise} show the results for
our full mock data ensemble. Each subplot corresponds to a different mock
galaxy, as marked. We show the fractional error of the mass distribution for
each of the test configurations with (red) and without (black) stellar mass. In
\figref{main results} we define the error
%
\begin{equation} \label{ferror R}
  f_R = \frac {\sum_i \left|M(i) - \tilde M(i)\right| } {\sum \tilde M(i)}
\end{equation}
%
based on the mass $M(i)$ of each pixel ring $i$ and the mass $\tilde M$ from the mock galaxy. 
In \figref{main results pixel-wise} the error is defined over all the pixels $\vec\theta$:
%
\begin{equation} \label{ferror theta}
f_\theta = \frac {\sum_{\vec\theta} \left|M(\vec\theta) - \tilde M(\vec\theta)\right| } {\sum_{\vec\theta} \tilde M(\vec\theta)}
\end{equation}
For both $f_R$ and $f_\theta$, we only consider mass up to one pixel length
passed the outermost image, since there is no longer any lensing information
beyond that point. This means we typically use 8 bins, linearly spaced,
ignoring the outermost 3 bins.  The spacing changes, however, at the border
between the high resolution region in the middle.

>>>>

In this paper we demonstrate which combination of lensing, stellar mass, and
stellar kinematic constraints best constrain the projected mass profile and
shape of several gravitational lens produced from mock data. In particular,
we introduce the idea of using estimates of the stellar mas content to place
lower bounds on the lensing mass.
We also introduce a new lens modeling framework  -- \Glass\ (Gravitational
Lensing AnalysiS Software). The compute-intensive core of \Glass\ is a parallel
implementation of the high-dimensional sampling algorithm from
\cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has extensible code (in
software terminology, a framework) in Python, encoding the lens equation, a
basis and priors for the mass distribution, and various post-processing and
visualisation modules.

\Glass\ is described in \secref{sec:glass}.  In \secref{sec:theory}, we review
the key elements of lensing theory, stellar population synthesis, and stellar
dynamics we will need. In \secref{sec:mockdata}, we describe our mock data. In
\secref{sec:results}, we present our results from applying \Glass\ to these
mock data. Finally, in \secref{sec:conclusions} we present our conclusions.

@@

Unfortunately, the lens equation and the arrival times alone are not enough to
form a closed volume in the solution space. We have not imposed any strong
constraints on the density even so much as to ensure that it is non-negative
everywhere. We must therefore provide additional linear constraints so that we
only sample physical models. Below we list typical constraints that \Glass\
provides. The explicit implementation for these priors has
been explained in \cite{2008ApJ...679...17C}, but can be summarised as
%
\begin{enumerate}
\item The density must be non-negative everywhere.
\item The azimuthally averaged density profile must have a slope everywhere $\le 0$.
\item Image parity is enforced.
\item The local gradient everywhere must point within $45^{\circ}$ of the center.
\item Density variations must be smooth.
\item The density is inversion symmetric.
\end{enumerate}

The first three of these priors are very important for creating physically
meaningful lenses. Prior (ii) is necessary to prevent rings from forming where
the inner region is less dense. This would be unusual for a galaxy or a
cluster. Prior (iv) can be relaxed somewhat by increasing the angle. The
smoothness prior (v) requires pixel values to be less than a constant factor of
the average neighbor value. This can be relaxed by increasing the constant or
completely disabled. For this paper, we have chosen \emph{not} to use this
prior, and therefore allow for the presence of small density substructure.
Prior (vi) enforces a radial symmetry that requires radially opposed pixels
(i.e., on opposite sides of the center) to be equal. This prior is particularly
important for modelling doubles which must be symmetric. If a lens is suspected
to be symmetric this can also significantly improve the modelling time for
quads.

While most of these priors have been selected for modelling galaxies, there
are several ways to address clusters or allow for the presence of substructure.
As previously described, one can add any number of additional analytic 
potentials inside or outside the main modelling pixelation area. For instance,
in a cluster, one can explicitly account for galaxies with power law potentials.
Furthermore, Priors (iv) and (v) and be relaxed or disabled. We have perfomed
preliminary testing of these ideas several common clusters and are encouraged
by the results. We will reserve a full analysis for a future publication.

In the simplest form, a single model for a lens is a tuple $\M = (\vec\kappa,
\vec\beta, \gamma_1, \gamma_2)$. A single model represents a single point in
the solution space polytope. Using the MCMC sampling strategy described in
\cite{2012MNRAS.425.3077L} we uniformly sample this space. Collectively, the
sampled models are referred to as an ensemble $\E = \{\M_i\}$, where we usually
generate $|\E| \sim 1000$ models. One can choose to further process these
models to impose further priors that may be difficult to enforce during the
modeling process. For instance, non-linear constraints, or simply filtering
of models that do not meet some criteria can be excluded, or weighted against.
For our current analysis, we do not exclude any models.

The time to generate the model ensemble is mostly a function of the size of the
parameter space. The MCMC algorithm has a ``warm-up'' phase where it estimates
the size and shape of each dimension in the solution space.  Once this has been
completed, the models are sampled very quickly. In fact, there is little
difference between generating 1,000 or 10,000 models. For our problems, the
typical ``warm-up'' time was about 4s, and the modelling time was 20s using a
parallel shared-memory machine with 40 cores.

@@

\figref{main results} and \figref{main results pixel-wise} show the results for
our full mock data ensemble. Each subplot corresponds to a different mock
galaxy, as marked. We show the fractional error of the mass distribution for
each of the test configurations with (red) and without (black) stellar mass. In
\figref{main results} we define the error
%
\begin{equation} \label{ferror R}
  f_R = \frac {\sum_i \left|M(i) - \tilde M(i)\right| } {\sum \tilde M(i)}
\end{equation}
%
based on the mass $M(i)$ of each pixel ring $i$ and the mass $\tilde M$ from the mock galaxy. 
In \figref{main results pixel-wise} the error is defined over all the pixels $\vec\theta$:
%
\begin{equation} \label{ferror theta}
f_\theta = \frac {\sum_{\vec\theta} \left|M(\vec\theta) - \tilde M(\vec\theta)\right| } {\sum_{\vec\theta} \tilde M(\vec\theta)}
\end{equation}
%
For both $f_R$ and $f_\theta$, we only consider mass up to one pixel length
passed the outermost image, since there is no longer any lensing information
beyond that point. This means we typically use 8 bins, linearly spaced,
ignoring the outermost 3 bins.  The spacing changes, however, at the border
between the high resolution region in the middle.
